\input{../templateExo.tex}

\begin{document}
\hrulefill

\begin{minipage}{0.33\textwidth}
\textsc{Math 471}
\end{minipage} \hfill 
\begin{minipage}{0.32\textwidth}
\centering
\textsc{Problems Set} \\
Appendix L
\end{minipage}
 \hfill 
 \begin{minipage}{0.33\textwidth}
 \flushright \textsc{Fall 2023}
 \end{minipage}

\hrulefill

\setcounter{section}{12}

 \subsection[~~Mathematical Statements]{Mathematical Statements}

 \begin{problem}
 Are the following sentences statements? If it is, say if it is true or false. If it is not, explain briefly why.
 	\begin{enumerate}[label=\alph*)]
 	\item $|-12| = -12$.
 	\item $x < 0$.
 	\item Is that an odd integer?
 	\item If $a= 2$ and $b = 4$, then $a + b = 6$.
 	\end{enumerate}
 \end{problem}

 \subsection[~~Logic and Mathematical Language]{Logic and Mathematical Language}

 \begin{problem}
Give the converse and the contrapositive of the following conditional statements.
 	\begin{enumerate}[label=\alph*)]
 		\item If it is Saturday, then Angela sleeps in.
 		\item If it rains outside, then I use my umbrella.
 		\item If I surf, then the waves are bigger than 4 feet high.
 	\end{enumerate}
 \end{problem}

 \begin{problem}
 Write useful negations of the following statements in English. You can use symbols to simplify the statement.
 	\begin{enumerate}[label=\alph*)]
 		\item It is raining and Charlie is cold.
 		\item If it is raining, then Charlie is cold.
 		\item For every real number $x$, there exists a real number $y$ such that $x  +y = 0$.
 		\item $|a| > 0$ if and only if $a \neq 0$.
 	\end{enumerate}
 \end{problem}

 \begin{problem}
 	\begin{enumerate}[label=\alph*)]
 		\item  By constructing the truth table of $P \Rightarrow Q$ and $Q \Rightarrow P$, show when a conditional statement and its converse do not have the same truth values.
 		\item By constructing the truth table of $P \Rightarrow Q$ and $(\neg Q) \Rightarrow (\neg P)$, show a conditional statement and its contrapositive always have the same truth values.
 	\end{enumerate}
 \end{problem}

 \subsection[~~Methods of Proof]{Methods of Proof}

 \begin{problem}
 Suppose that $a$ and $b$ are integers. Prove each of the following.
 	\begin{enumerate}[label=\alph*)]
 		\item If $a$ and $b$ are both odd, then $a + b$ is even.
 		\item If $a$ is even and $b$ is odd, then $a + b$ is odd.
 	\end{enumerate}
 \end{problem}

 \begin{problem}
 A rational number is a number $q$ that can be put in the form of a fraction, that is there exist two integers $n$ and $m$ such that $q = n / m$. Show that $\sqrt{2}$ is not rational.
 \end{problem}

 \begin{problem}
 Prove that there exist integers $m$ and $n$ such that $2m + 3n = 12$.
 \end{problem}

\end{document}